\section{Synopsis}
\label{Synopsis}

Traditional Bayesian epistemology used to include the
requirement for a rational agent to hold a sharp
credence function. It has recently become popular to
drop this requirement. There are now Bayesian theories
which permit a rational agent to hold indeterminate
credal states (from now on instates) based on
incomplete or ambiguous evidence. I will refer to
Bayesians who continue to adhere to the classical
theory of sharp credences for rational agents as
\qnull{Laplaceans} (e.g.\ Adam Elga and Roger White). I
will refer to Bayesians who do not believe in the
requirement for a rational agent to hold sharp
credences as \qnull{Booleans} (e.g.\ Peter Walley and
James Joyce; see Boole, 1854, chapters 16--21, for
alternative methods to the ones suggested by Laplace
which result in imprecise epistemic probabilities).

After describing the appeal of indeterminacy and
showing how contemporary Laplacean objections fail, I
will point to more serious failings of indeterminacy in
semantic terms and show how a proper semantics of not
knowing, which we could also call a semantics of
partial belief, solves the problems for sharp credences
that Booleans address by introducing instates. There is
a sense in which, by linking knowledge of chances to
its reflection in credences, Booleans seek to reconcile
traditional knowledge epistemology concerned with full
belief on the one hand and formal epistemology
concerned with partial belief on the other hand. There
are other more recent reconciliation projects (see
Spohn, 2012; and Moss, 2013), but if my paper is
correct then the Boolean approach will not contribute
to this reconciliation because it mixes full belief and
partial belief metaphors in ways that are semantically
problematic.

A sharp credence, as much as the term suggests
precision and a measure of certainty, is a
representation of an epistemic state characterized by
uncertainty and lack of information. Importantly, it
does not represent the evidence which informs the
epistemic state, and it makes no claim of such a
representation. Instates are often lauded as doing much
better representing uncertainty together with the
evidence that constrains it, but they can no more give
an adequate representation of evidence than sharp
credences. This paper claims that it is legitimate to
have a $0.5$ sharp credence in heads for a coin of
whose bias we are completely ignorant; for a coin whose
fairness is supported by a lot of evidence; and even
for a coin about whose bias we know to be either 1/3 or
2/3 for heads.

One potential Boolean claim is that agents who use
instates do better than Laplaceans when they bet on the
truth of events for which they have varying degrees of
evidence. Walley gives an example where a Laplacean
does much worse at predicting soccer games than Boolean
peers. I show that the result is due to an unreasonable
restriction on the betting behaviour of the Laplacean.
Once this restriction is lifted, Laplaceans do just as
well as Booleans, except that they are not tangled in
the semantic problem of the double task, where instates
are supposed to reflect both the uncertainty of an
agent and other properties of her evidence.

I will present several examples where this double task
stretches instates to the limits of plausibility, for
example when they need to aggregate expert opinion or
account for dilation. Joyce's idea that credences can
represent balance, weight, and specificity of the
evidence is inconsistent with the use of indeterminacy
(and Joyce himself, in response to the dilation
problem, gives the argument why this is the case). The
implicit Boolean claim that certain properties of the
evidence (its ambiguity, its completeness, conflicts
within it) can be recovered from instates is
inconsistent with an effective Boolean answer to the
dilation problem.

The Laplacean approach of assigning subjective
probabilities to partitions of the event space (e.g.\
objective chances) and then aggregating them by David
Lewis's summation formula into a single precise
credence function is semantically tidy and shares many
of the formal virtues of Boolean theories. If a bad
taste about numerical precision in our fuzzy, nebulous
world lingers, I can point to philosophical projects in
other domains where the concepts we use are sharply
bounded, even though our ability to conceive of those
sharp boundaries or know them is limited, for example
Timothy Williamson's work on vagueness and knowledge as
a mental state.

\section{Motivation for Instates}
\label{MotivationForIndeterminateCredalStates}

Booleans claim that it is rational for an agent to hold
instates (indeterminate credal states) in addition to
sharp credences, representing uncertainty. Booleans
defend Bayesian epistemology in all other respects,
proposing that it is better off without the requirement
for sharp credences. Laplaceans, by contrast, require
sharp credences for rational agents. Indeterminacy is
forcefully motivated in the Boolean literature. The
reader will see (a) the appeal of the Boolean approach,
(b) the insufficiency of the critical response, and (c)
the need for careful articulation of the Laplacean
approach that can address the concerns which motivate
some to resort to indeterminacy. Finally, (d) the
undesirable semantics of the Boolean approach,
documented both conceptually and by example, will lead
to the conclusion that a Laplacean approach is the more
promising alternative.

Two compelling reasons for the Boolean approach are
that instates sometimes represent evidence better than
sharp credences, to the point where sharp credences
misrepresent the evidence; and that the preference
structure of a rational agent may be incomplete so that
representation theorems do not yield sharp credences to
represent such incomplete structures. A third reason is
poorly documented: the Boolean rational agent may do
better accepting advantageous bets than the agent who
on principle rejects instates. Walley conducted an
experiment in which Boolean participants did
significantly better than Laplacean participants,
betting on soccer games played in the Soccer World Cup
1982 in Spain (see \scite{7}{walley91}{}, Appendix
F).\tbd{check reference} I replicated the experiment
using two computer players with rudimentary artificial
intelligence and made them specify betting parameters
(previsions) for games played in the Soccer World Cup
2014 in Brazil. I used the Poisson distribution (which
is an excellent predictor for the outcome of soccer
matches) and the FIFA ranking to simulate millions of
counterfactual World Cup results and their associated
bets, using Walley's evaluation method. The Boolean
player had a slight but systematic advantage. In
section \ref{WalleysWorldCupWoes}, I will provide an
explanation and show how it undermines any support the
experiment might give to the Boolean position.

\section{Dilation}
\label{Dilation}

Roger White introduces a problem for instates claiming
that they lead to unacceptable doxastic scenarios
involving dilation. In dilation, there is a widening of
the instate upon updating instead of a narrowing.
Sometimes this widening is extreme, from a maximally
precise credal state to a vacuous instate, based on
little information (see \scite{7}{white10}{}, for
examples). White's objection, even though I think it
fails as it stands, triggers semantic concessions on
the part of Booleans defending their position. They
will be important to my semantic criticism of
indeterminacy.

In Joyce's response to White, two semantic concessions
to White show why the dilated instates give us the
right result. I agree with Joyce: dilation is what you
would expect (1) if credences do not adequately
represent evidence (the same instate can reflect
different evidential situations); and (2) if instates
do not reflect knowledge claims about objective chances
(Joyce rejects White's Chance Grounding Thesis CGT, see
\scite{8}{joyce10}{289}). Dilation from a sharp $0.5$
to an indeterminate $\{0.01,0.99\}$ or $[0.01,0.99]$
(depending on whether convexity is required) is
unproblematic in the following example, although the
example already prefigures that there is something odd
about the Boolean semantic approach. The example
licences a 99:1 bet for one of the colours, but this is
a problem that arises out of Boolean semantics without
dilation, which we will address again in Example
\ref{monkey}.

\addtocounter{expls}{1}

\begin{quotex} \textbf{Example \arabic{expls}:
    \addtocounter{expls}{1} Dilating Urns.} You draw
  from an urn with 200 balls (100 red, 100 black) and
  receive the information that the urn actually had two
  chambers, one with 99 red balls and 1 black ball, the
  other with 1 red ball and 99 black balls.
\end{quotex}

If one were to be committed to the principle of
regularity, that all states of the world considered
possible have positive probability (for a defence see
\scite{7}{edwardsetal63}{}); and to the solution of
Henry Kyburg's lottery paradox, that what is rationally
accepted should have probability 1 (for a defence of
this principle see \scite{7}{douvenwilliamson06}{});
and the CGT, that one's spread of credence should cover
the range of possible chance hypotheses left open by
evidence (implied by much of Boolean literature); then
one's instate is always vacuous. Booleans must deny at
least one of the premises to avoid the conclusion.
Joyce denies the CGT, but then he continues to make
implicit use of it when he repeatedly complains that
sharp credences \qeins{ignore a vast number of
  possibilities that are consistent with [the]
  evidence} (for example in \scite{8}{joyce05}{170}).

When updating dilates the credal state, it appears that
the prior credal state was in some sense incorrect and
did not properly reflect the state of the world, even
though it properly reflected the epistemic state of the
agent. Such a divergence between the proper reflections
of epistemic state and state of the world undermines
the subjective interpretation of probabilities at the
heart of Bayesian epistemology. An instate semantically
blurs the line between traditional full belief
epistemology and formal partial belief epistemology
(for semantically more intelligible attempts at
reconciliation between the two see \scite{7}{moss13}{},
although Moss is committed to the Boolean approach,
which may weaken her case; and \scite{7}{spohn12}{},
especially chapter 10). White's objection fails,
however, because dilation for indeterminate credences
is in principle not any more surprising than a piece of
information that increases the Shannon entropy of a
sharp credence. It is true for both sharp and
indeterminate credences that information can make us
less certain about things.

\section{Semantics of Partial Belief}
\label{SemanticsOfPartialBelief}
Central to my project is the proper semantic
distinction between evidence, information, and partial
belief. Both sharp credences and instates, within a
Bayesian framework, try to represent the uncertainty of
an agent with respect to the truth of a proposition.
Instates have a greater ambition and therefore a double
task: they also claim to represent properties of the
evidence, such as its weight, conflict between its
constituents, or its ambiguity.

The Laplacean approach uses partial beliefs about how a
parameter is distributed and then applies Lewis'
summation formula (see \scite{8}{lewis81}{266f}) to
integrate over them and condense them to a sharp
credence. Walley comments on this \qeins{reduction} in
his section on Bayesian second order probabilities (see
\scite{8}{walley91}{258f}), but he mistakenly
represents the Laplacean approach as a second order
approach, as if the probability distributions that are
summarized by Lewis' formula are of the same kind as
the resulting credences. They are not. They are
objective chances or other partitions of the event
space and the subjective probabilities that are
associated with them. It is the Boolean approach which
has elements of a second order approach and thus makes
itself vulnerable to regress problems by adding another
dimension of uncertainty to a parameter (the credence)
which already represents uncertainty.

One of Joyce's complaints is that a sharp credence of
$0.5$ for a coin contains too much information if there
is little or no evidence that the coin is fair. This
complaint, of course, is only effective if we make a
credence say something about the evidence. Joyce
himself, however, admits that instates cannot
adequately represent the evidence without violating the
reflection principle due to White's dilation problem.
He is quite clear that the same instate can represent
different evidential scenarios (see, for example,
\scite{8}{joyce10}{302}).

Sharp credences have one task: to represent epistemic
uncertainty and serve as a tool for updating,
inference, and decision-making. They cannot fulfill
this task without continued reference to the evidence
which operates in the background. To use an analogy,
credences are not sufficient statistics with respect to
updating, inference, and decision-making. What is
remarkable about Joyce's response to White's dilation
problem is that Joyce recognizes that instates are not
sufficient statistics either. But this means that they
fail at the double task which has been imposed on them.
In the following, I will provide a few examples
illustrating this failure.

\begin{quotex} 
  \textbf{Example \arabic{expls}:
    \addtocounter{expls}{1} Aggregating Expert
    Opinion.} You have no information whether it will
  rain tomorrow ($R$) or not except the predictions of
  two weather forecasters. One of them forecasts 0.3 on
  channel GPY, the other 0.6 on channel QCT. You
  consider the QCT forecaster to be significantly more
  reliable, based on past experience.
\end{quotex}

An instate corresponding to this situation may be
$[0.3,0.6]$ (see \scite{8}{walley91}{214}), but it will
have a difficult time representing the difference in
reliability of the experts. A sharp credence of
$P(R)=0.55$, for example, does the right thing. Beliefs
about objective chances make little sense in many
situations where we have credences, since it is
doubtful even in the case of rain tomorrow that there
is an urn of nature from which balls are drawn (see
\scite{8}{white10}{171}). What is really at play is a
complex interaction between epistemic states (for
example, experts evaluating meteorological data) and
the evidence which influences them.

\begin{quotex}
  \textbf{Example \arabic{expls}:
    \addtocounter{expls}{1}\label{monkey} Precise
    Credences.} Your sharp credence for rain tomorrow,
  based on the expert opinion of channel GPY and
  channel QCT (you have no other information) is
  $0.55$. Is it reasonable, considering how little
  evidence you have, to reject the belief that the
  chance of rain tomorrow is $0.54$ or $0.56$; or to
  prefer a 54.9 cent bet on rain to a 45.1 cent bet on
  no rain?
\end{quotex}

The first question is confused in instructive ways. A
sharp credence rejects no hypothesis about objective
chances (unlike an instate). It often has a subjective
probability distribution operating in the background,
over which it integrates to yield the sharp credence.
The subjective probability distribution is condensed by
Lewis' summation formula to a sharp credence, without
being reduced to it. Lewis' 1981 paper \qeins{A
  Subjectivist's Guide to Objective Chance} addresses
the question of the relationship between credence,
subjective probability, and objective chance. The
Laplacean properly separates the semantic dimensions,
without using second-order probabilities. Formal
partial belief epistemology deals with sharp credences
and how they represent uncertainty and serve as a tool
in inference, updating, and decision making; Bayesian
epistemology provides the interpretation of
probabilities; and Lewis' Humean speculations and his
interpretation of the principal principle cover the
relationship between subjective probabilities and
objective chance.

The second question is also instructive: why would we
prefer a 54.9 cent bet on rain to a 45.1 cent bet on no
rain, given that we do not possess the power of
discrimination between these two bets? The answer to
this question ties in with the issue of incomplete
preference structure referenced above as a motivation
for instates. Via representation theorems, preferences
may conceptually precede an agent's probability and
utility functions, but that does not mean that we
cannot inform the axioms we use for a rational agent's
preferences by undesirable consequences downstream.
Completeness may sound like an unreasonable imposition
at the outset, but if incompleteness has unwelcome
semantic consequences for credences, it is not
illegitimate to revisit the issue. Timothy Williamson
goes through this exercise with vague concepts, showing
that all upstream logical solutions to the problem fail
and that it has to be solved downstream with an
epistemic solution (see \scite{7}{williamson96}{}).

Vague concepts, like sharp credences, are sharply
bounded, but not in a way that is luminous to the agent
(for anti-luminosity see chapter 4 in
\scite{7}{williamson00}{}). Anti-luminosity answers the
original question: the rational agent prefers the 54.9
cent bet on rain to a 45.1 cent bet on no rain based on
her sharp credence without being in a position to have
this preference necessarily or have it based on
physical or psychological ability (for the analogous
claim about knowledge see \scite{8}{williamson00}{95}).

In a way, advocates of indeterminacy have solved this
problem for us. There is strong agreement among most of
them that the issue of indeterminacy for credences is
not an issue of elicitation. The appeal of preferences
is that we can elicit them more easily than assessments
of probability and utility functions. The indeterminacy
issue has been raised to the probability level (or
moved downstream) by indeterminacy advocates themselves
who feel justifiably uncomfortable with an
interpretation of their theory in behaviourist terms.
So it shall be solved there, and this paper makes an
appeal to reject indeterminacy on this level. Isaac
Levi seems to agree with me on this point: when he
talks about indeterminacy, it proceeds from the level
of probability judgment to preferences, not the other
way around (see \scite{8}{levi81}{533}).

\begin{quotex}
  \textbf{Example \arabic{expls}:
    \addtocounter{expls}{1} Jaynes' Monkeys.} Let urn
  $A$ contain four balls, two red and two black. A
  monkey randomly fills urn $B$ from urn $A$ with two
  balls. We draw from urn $B$. (See
  \scite{8}{jaynesbretthorst03}{160}.)
\end{quotex}

The sharp credence of drawing a red ball is $0.5$,
following Lewis' summation formula for the different
combinations of balls in urn $B$. This solution is more
intuitive in terms of further inference, decision
making, and betting behaviour than an instate of
$\{0,1/2,1\}$ or $[0,1]$ (depending on the convexity
requirement), since the instate would licence an
exorbitant bet in favour of one colour.

\begin{quotex}
  \textbf{Example \arabic{expls}:
    \addtocounter{expls}{1} Three Prisoners.} The Three
  Prisoners problem is well-documented (see
  \scite{8}{mosteller87}{28}). Let the three prisoners
  be $X_{1},X_{2},X_{3}$, and the warden tell $X_{1}$
  that $X_{3}$ will be executed.
\end{quotex}

Walley maintains that for the Monty Hall problem and
the Three Prisoners problem, the probabilities of a
rational agent should dilate rather than settle on the
commonly accepted solutions. For the Three Prisoners
problem, there is a compelling case for standard
conditioning and the result that the chances of pardon
for prisoner $X_{1}$ are unchanged after the update
(see \scite{8}{lukits14}{1421f}). Walley's dilated
solution would give prisoner $X_{1}$ hope on the
doubtful possibility (and unfounded assumption) that
the warden might prefer to provide $X_{3}$'s name in
case prisoner $X_{1}$ was pardoned.

Booleans charge that sharp credences often reflect
independence of variables where such independence is
unwarranted. Booleans prefer to dilate over the
possible dependence relationships (independence
included). White's dilation problem is an instance of
this. The fallacy in the argument for instates,
illustrated by the Three Prisoners problem, is that the
probabilistic independence of sharp credences does not
imply independence of variables (the converse is
correct), but only that it is unknown whether there is
dependence, and if yes, whether it is correlation or
inverse correlation.

\begin{quotex}
  \textbf{Example \arabic{expls}:
    \addtocounter{expls}{1} Wagner's Linguist.} A
  linguist hears the utterance of a native and
  concludes that the native cannot be part of certain
  population groups, depending on what the utterance
  means. The linguist is uncertain between some options
  about the meaning of the utterance. (For full details
  see \scite{8}{wagner92}{252}; and
  \scite{8}{spohn12}{197}.)
\end{quotex}

The mathematician Carl Wagner proposed a natural
generalization of Jeffrey Conditioning for his Linguist
example (see \scite{7}{wagner92}{}). Since the
principle of maximum entropy is already a
generalization of Jeffrey Conditioning, the question
naturally arises whether the two generalizations agree.
Wagner makes the case that they do not agree and
deduces that the principle of maximum entropy is
sometimes an inappropriate updating mechanism, in line
with many earlier criticisms of the principle of
maximum entropy (see van Fraassen, 1981;
\scite{7}{shimony85}{}; \scite{7}{skyrms87updating}{};
and, later on, \scite{7}{grovehalpern97}{}).

What is interesting about this case is that Wagner uses
instates for his deduction, so that even if you agree
with his natural generalization of Jeffrey Conditioning
(which I find plausible), the inconsistency with the
principle of maximum entropy can only be inferred
assuming instates. On the assumption of sharp credences
Wagner's generalization of Jeffrey conditioning
perfectly accords with the principle of maximum
entropy. Wagner's argument, instead of undermining the
principle of maximum entropy, just shows that instates
are as wedded to rejecting the claims of the principle
of maximum entropy as the principle of maximum entropy
is wedded to sharp credences.

\section{Evidence Differentials and Cushioning
  Credences}
\label{WalleysWorldCupWoes}

After I found out that agents with instates do better
betting on soccer games (see section
\ref{MotivationForIndeterminateCredalStates}), I let
player $X$ (who uses sharp credences) and player $Y$
(who uses instates) play a more basic betting game and
used Walley's pay off scheme (see
\scite{8}{walley91}{632}) to settle the bets. The
simulation results show that player $Y$ does better for
$n>2$, while player $X$ does better for $n=2$. A
defence of sharp credences for rational agents needs to
have an explanation why, for $n>2$, player $Y$ does
better. We will call it partial belief cushioning,
which prevents a uniform distribution of bets over
cases where there is an evidence differential and
enables greater density of bets where evidence is more
solid.

Compulsory betting is a problem for instates, as
Booleans have to find a way to decide without
necessarily receiving instructions from the credal
state. Booleans have addressed this point extensively
(see for example \scite{8}{joyce10}{311ff}; for an
opponent's view of this see \scite{8}{elga10}{6ff}).
The problem for sharp credences arises when bets are
noncompulsory, for then the data above suggests that
agents holding instates sometimes do better. Where
evidence is thin, an agent with sharp credences must
cushion her credences and is better off by rejecting
bets that look attractive in terms of her partial
beliefs. If an agent does not cushion her partial
beliefs (whether they are sharp or indeterminate), she
will incur a loss in the long run. Since cushioning is
permitted in Walley's experimental setup (the bets are
noncompulsory), Laplacean agents should also have
access to it and then no longer do worse than Boolean
agents.

The rational agent with a sharp credence has resources
at her disposal to use just as much differentiation
with respect to accepting and rejecting bets as the
agent with instates. Often (if she is able to and
especially if the bets are offered to her by a
better-informed agent), she will reject both of two
complementary bets, even when they are fair. In 2013,
the U.S. Commodity Futures Trading Commission shut down
the online prediction market Intrade, even though it
was offering fair bets, because of expert arbitrageurs.

Any advantage that the agent with an instate has can be
counteracted based on distributions over partial
beliefs with respect to all possibilities. The agent
with instates suffers from a semantic mixing of
metaphors between evidential and epistemic dimensions
that puts her at a real disadvantage in terms of
understanding the sources and consequences of her
knowledge and her uncertainties.
