Caticha, Ariel, and Adom Giffin. ``Updating Probabilities.'' \emph{In MaxEnt 2006, the 26th International Workshop on Bayesian Inference and Maximum Entropy Methods}. 2006.

Cover, T.M., and J.A. Thomas. \emph{Elements of Information Theory}, volume 6. Hoboken, NJ: Wiley, 2006.

Dempster, Arthur. ``Upper and Lower Probabilities Induced by a Multivalued Mapping.'' \emph{The Annals of Mathematical Statistics} 38, 2: (1967) 325--339.

Friedman, Kenneth, and Abner Shimony. ``Jaynes's Maximum Entropy Prescription and Probability Theory.'' \emph{Journal of Statistical Physics} 3, 4: (1971) 381--384.

Guia{\c{s}}u, Silviu. \emph{Information Theory with Application}. New York: McGraw-Hill, 1977.

Howson, Colin, and Allan Franklin. ``Bayesian Conditionalization and Probability Kinematics.'' \emph{The British Journal for the Philosophy of Science} 45, 2: (1994) 451--466.

Jaynes, E.T. ``Where Do We Stand on Maximum Entropy.'' In \emph{The Maximum Entropy Formalism,} edited by R.D. Levine, and M. Tribus. Cambridge, MA: MIT, 1978, 15--118.

Jeffrey, Richard. \emph{The Logic of Decision}. New York, NY: Gordon and Breach, 1965.

Majern{\'\i}k, Vladim{\'\i}r. ``Marginal Probability Distribution Determined by the Maximum Entropy Method.'' \emph{Reports on Mathematical Physics} 45, 2: (2000) 171--181.

Spohn, Wolfgang. \emph{The Laws of Belief: Ranking Theory and Its Philosophical Applications}. Oxford, 2012.

Teller, Paul. ``Conditionalization and Observation.'' \emph{Synthese} 26, 2: (1973) 218--258.

Wagner, Carl. ``Generalized Probability Kinematics.'' \emph{Erkenntnis} 36, 2: (1992) 245--257.

Wagner, Carl. ``Probability Kinematics and Commutativity.'' \emph{Philosophy of Science} 69, 2: (2002) 266--278.
