% This is NOT the official version, but a seriously cut down version for APA
% Pacific 2014. If you make corrections here, they must be made in sonk.tex as
% well if they are meant to endure.

% TBD: 

% (1) What exactly is player $X$'s evidence differential compared to
% player $Y$ -- the analytic expression may help.

% (2) Talk about Second Order probabilities, especially with reference
% to hierarchical nesting, see Second order in collator, 

\documentclass[11pt]{article}
\usepackage{october}
% For BJPS
% \hyphenpenalty=10000
% \hbadness=10000

\begin{document}
% For BJPS
% \raggedright
% \doublespacing

\title{Semantics of Not Knowing}
\author{Stefan Lukits}
\date{}
\maketitle
% \doublespacing

% \begin{abstract} 
%   {\noindent}
% \end{abstract}

\section{Introduction}
\label{Introduction}

See 750 word synopsis.

\section{Motivation for Indeterminate Credal States} 
\label{MotivationForIndeterminateCredal States}

Booleans claim that it is rational for an agent to hold an
indeterminate credal state towards an event. The indeterminate credal
state represents uncertainty about the truth value of the event and
informs decision-making, inference, and betting behaviour. Booleans
are Bayesians in all other respects and defend Bayesian epistemology,
proposing that it is better off without the requirement for sharp
credences. Laplaceans, by contrast, require sharp credences for
rational agents. We want to motivate indeterminacy, independent of how
they are elicited from the agent, as forcefully as possible so that
the reader will see (a) the appeal of such indeterminacy, (b) the
insufficiency of the critical response, and (c) the need for careful
articulation of the Laplacean approach that mandates a rational agent
to hold sharp credences together with an explanation of how Laplaceans
can address Finally, (d) the undesirable semantics of the Boolean
approach, documented both conceptually and in practice, will lead to
the conclusion that the Laplacean approach is desirable.

Let a \textsc{coin} be a Bernoulli generator that produces successes
and failures with probability $p$ for success, labeled $H$, and $1-p$
for failure, labeled $T$. Physical coins may serve as examples if we
are willing to set aside that most of them are approximately fair.
Imagine three \textsc{coin}s for which we have evidence that
\textsc{coin}$_{I}$ is fair, \textsc{coin}$_{II}$ has an unknown bias,
and \textsc{coin}$_{III}$ has as bias either $p=1/3$ or $p=2/3$. The
Laplacean approach, on the one hand, permits a sharp $0.5$ credence in
$H$ for a rational agent in all three cases. A Boolean approach, on
the other hand, sees the difference in the evidential situation
reflected in a rational agent's credal state and for example at least
permits, as credence in $H$, $\{x|x=0.5\}$ for \textsc{coin}$_{I}$,
$\{x|0\leq{}x\leq{}1\}$ for \textsc{coin}$_{II}$, and
$\{x|1/3\leq{}x\leq{}2/3\}$ or $\{1/3,2/3\}$ for
\textsc{coin}$_{III}$.

Here is a list of reasons why credal states for ideally
rational agents do not need to be representable by single probability
measures.

\begin{enumerate}[(A)]
\item By far the greatest emphasis in motivation for indeterminacy
  rests on lack of evidence or conflicting evidence and the assumption
  that single probability measures (sharp credences) do not represent
  such evidence as well as credal states composed by sets of
  probability measures (indeterminacy credal states).
\item The preference structure of a rational agent may be incomplete
  so that representation theorems do not yield single probability
  measures to represent such incomplete structures.
\item There are more technical and paper-specific reasons, such as
  Thomas Augustin's attempt to mediate between the minimax pessimism
  of objectivists and the Bayesian optimism of subjectivists using
  interval probability (see \scite{8}{augustin03}{35f}); Alan
  H{\'a}jek and Michael Smithson's belief that there may be
  objectively indeterminate chances in the physical world (see
  \scite{8}{hajeksmithson12}{33}); and Jake Chandler's claim that
  \qeins{the sharp model is at odds with a trio of plausible
    propositions regarding agnosticism} \scite{2}{chandler14}{4}.
\end{enumerate}

This paper mostly addresses (A), while taking (B) seriously as well and
pointing towards solutions for it.  I am leaving (C) to more specific
responses to the issues presented in the cited articles, and for the remainder
of this section I am adding a reason (D) that is poorly documented in the
literature.

One motivation for permitting a rational agent to have a range of probability
measures rather than mandating her to hold a single one as her credence is
that she would do better making decisions and accepting advantageous bets.
Upper and lower previsions, components of a plausible version of indeterminacy
(see \scite{7}{walley91}{}), have a behavioural interpretation as maximum
buying prices and minimum selling prices for gambles where two bets can both
be rejected if they offer a reward of \$1 for a price of $x$ in case of $H$
and a reward of \$1 for a price of $1-x$ in case of $T$ as long as $x$ falls
between the upper and lower prevision.

Walley conducted an experiment with 17 participants, who had various
levels of understanding Bayesian theory, asking them for upper and
lower probabilities entering into bets about the games played in the
Soccer World Cup 1982 in Spain. One of the participants, a
\qeins{dogmatic Bayesian lecturer} (see \scite{8}{walley91}{633}),
only used single sharp subjective probabilities, while the others used
intervals. The dogmatic Bayesian lecturer finished a distant last when
the bets were evaluated.

I was curious what a more rigorous treatment of the comparison between
linear previsions (when upper and lower previsions coincide) and
interval previsions would yield. I equipped two computer players with
some rudimentary artificial intelligence and made them specify
previsions for games played in the Soccer World Cup 2014 in Brazil,
ordering player $X$ to specify linear previsions and player $Y$ to
specify upper and lower previsions. I used the Poisson distribution
(which is an excellent predictor for the outcome of soccer matches)
and the FIFA ranking to simulate millions of counterfactual World Cup
results and their associated bets, using Walley's evaluation method.
Player $Y$ had a slight but systematic advantage. In section
\ref{WalleysWorldCupWoes}, I will provide an explanation and show how
it can support rejecting indeterminate credal states for rational
agents, counter-intuitive as that may sound.

\section{Dilation} 
\label{Dilation}

Roger White introduces a problem for indeterminate credal states,
which I will call White's dilation problem, claiming that
indeterminate credal states lead to unacceptable doxastic scenarios
involving dilation. In dilation, there is a widening of the
indeterminate credal state upon updating instead of a narrowing
(especially Walley reiterates that more information generally means
more precision, see \scite{8}{walley91}{34}, although he also
addresses dilation, see \scite{8}{walley91}{299}). Sometimes this
widening is extreme, from a maximally precise credal state to a
vacuous credal state, based on very little information (see
\scite{7}{white10}{} for examples). White's objection, even though I
think it fails as it stands, triggers semantic concessions on the part
of Booleans defending their position which will be important to my
semantic criticism of indeterminacy.

Joyce's response to White, for example, uses two lines of defence: the
results are not counter-intuitive upon closer inspection and Lewis'
indadmissibility criterion as well as two significant semantic
concessions to White show why the indeterminate credal states give us
the right result. This response articulates just the right response to
White: yes, dilation is what you would expect (1) if credences do not
represent evidence (the same indeterminate credal state can reflect
different evidential situations); and (2) if the CGT is not a
necessary consequence of the indeterminacy approach.

Based on Joyce's amendments to the Boolean approach (and his own
version of it in \scite{7}{joyce05}{}), I agree with Joyce in White's
own words that White's dilation problem \qeins{is not a reductio but a
  result} \scite{3}{white10}{177}. The possible dependence between
\textsc{coin}$_{a}$ and \textsc{coin}$_{b}$ and the unknown bias of
\textsc{coin}$_{b}$ dilate our credences in the result of tossing
\textsc{coin}$_{a}$. The same would happen if we were drawing from an
urn with 200 balls (100 red, 100 black) and received the information
that the urn actually had two chambers, one with 99 red balls and 1
black ball, the other with 1 red ball and 99 black balls. Our credence
would dilate, given this additional piece of information, from a sharp
$0.5$ to an indeterminate $[0.01,0.99]$ without much mystery (the
indeterminate credence would licence an unreasonable 99:1 bet, but
that is part of the semantic criticism below).

A sharp credence, on the one hand, constrains partial beliefs in
objective chances by Lewis' summation formula (which we will provide
in the next section). No objective chance is excluded by it and any
updating will merely change the partial beliefs, but no full beliefs.
Indeterminate credal states, on the other hand, by giving ranges of
acceptable objective chances suggest that there is a full belief that
the objective chance does not lie outside what is indicated by the
indeterminate credal state.

If one were to be committed to the principle of regularity, that all
states of the world considered possible have positive probability (for
a defence see \scite{7}{edwardsetal63}{}); and to the solution of
Henry Kyburg's lottery paradox, that what is rationally accepted
should have probability 1 (for a defence of this principle see
\scite{7}{douvenwilliamson06}{}); and to White's CGT, that one's
spread of credence should cover the range of possible chance
hypotheses left open by evidence; then one's indeterminate credal
state must always be vacuous. Booleans must deny at least one of the
premises to avoid the conclusion. Joyce, for example, denies White's
CGT (see \scite{8}{joyce10}{289}), but then he continues to make
implicit use of it when he repeatedly complains that sharp credences
\qeins{ignore a vast number of possibilities that are consistent with
  [the] evidence} (for example in \scite{2}{joyce05}{170}). 

When updating dilates the credal state, it appears that the prior
credal state was in some sense incorrect and did not properly reflect
the state of the world, even though it properly reflected the
epistemic state of the agent. Such a divergence between the proper
reflections of epistemic state and state of the world undermines the
subjective interpretation of probabilities at the heart of Bayesian
epistemology. Credences are not knowledge claims about the world, but
represent uncertainty in the agent's epistemic state. Yet an
indeterminate credal state blurs the line between the two (for two
rigorous attempts to reconcile traditional full belief epistemology
and formal partial belief epistemology see \scite{7}{moss13}{}; and
\scite{7}{spohn12}{}, especially chapter 10).

White's objection fails because dilation for indeterminate credences
is in principle not any more surprising than a piece of information
that increases the Shannon entropy of a sharp credence. It is true for
both sharp and indeterminate credences that information can make us
less certain about things. Once Booleans have brought their house in
order to accommodate White's objection, however, they open wide the
door on semantic problems. We finally get to inquire what kind of
coherence there is in defending indeterminacy when it neither fulfills
the promise of representing evidence nor the promise of reconciling
traditional full belief \qnull{knowledge} epistemology and Bayesian
partial belief epistemology as outlined in the CGT, but only adds
another hierarchical layer of uncertainty to a numerical quantity (a
sharp credence) whose job it already is to represent uncertainty. We
will turn to these semantic considerations now, show how they display
the virtues of sharp credences in responding to the forceful
motivations for indeterminate credal states while making those
indeterminate credal states look semantically otiose. Then we will
show in the last section how sharp credences have an elegant solution
for being outperformed by indeterminate credal states in betting
scenarios, and there we will rest our case.

\section{Semantics of Partial Belief} 
\label{SemanticsOfPartialBelief}

At the heart of my project is a proper semantic distinction between
evidence, information, and partial belief. Both sharp credences and
indeterminate credal states, within a Bayesian framework, try to
represent the uncertainty of an agent with respect to the truth of a
proposition. Indeterminate credal states have a greater ambition: they
also claim to represent properties of the evidence, such as its
weight, conflict between its constituents, or its ambiguity. 

Indeterminate credal states are suggestive of a measurement that wants
to represent numerically the mass of an object and then also make
claims about its density. With sharp credences, the semantic roles of
evidence, information, and uncertainty are appropriately
differentiated. Rational decision-making, inference, and betting
behaviour are based on sharp credences together with the evidence that
is at its foundation. Information represents evidence, and sharp
credences represent uncertainty. Measurement is in any case a
misleading analogy for credences. Measurements are always imprecise.
Epistemic credences, however, are not measurements, especially not of
objective chances. They represent uncertainty. They are more like
logical truth values than they are like measurements.
\scite{8}{levi85}{407}, makes this category mistake, whereas
\scite{8}{walley91}{249} is clear on the difference.

It is important not to confuse the claim that it is reasonable to hold
both $X$ and $Y$ with the claim that it is reasonable to hold either
$X$ (without $Y$) or $Y$ (without $X$). It is the reasonableness of
holding $X$ and $Y$ concurrently that is controversial, not the
reasonableness of holding $Y$ (without holding $X$) when it is
reasonable to hold $X$. We will later talk about anti-luminosity, the
fact that a rational agent may not be able to distinguish
psychologically between a 54.9 cent bet on an event and a 45.1 bet on
its negation. She must reject one of them not to incur sure loss, so
proponents of indeterminacy suggest that she choose one of them freely
without being constrained by her credal state or reject both of them.
I claim that a sharp credence will make a recommendation between the
two so that only one of the bets is rational given her particular
credence, but that does not mean that another sharp credence which
would give a different recommendation may not also be rational for her
to have. 

There are two approaches you can take towards the indeterminacy
question, both of them involving two levels. On the one hand, you can
have partial beliefs about how a parameter is distributed and then use
Lewis' summation formula (see \scite{8}{lewis81}{266f}) to integrate
over them and condense them to a sharp credence. Walley comments on
this \qeins{reduction} in his section on Bayesian second order
probabilities (see \scite{8}{walley91}{258f}), but he mistakenly
represents the Laplacean approach as a second order approach, as if
the probability distributions that are summarized by Lewis' formula
are of the same kind as the resulting credences. They are not. They
are objective chances or other partitions of the event space and the
subjective probabilities that are associated with them. It is the
Boolean approach which has elements of a second order approach and
thus makes itself vulnerable to regress problems by adding another
dimension of uncertainty to a parameter (the credence) which already
represents uncertainty.

On the other hand, you can try to represent your uncertainty about the
distribution of the parameter by an indeterminate credal state. I want
to show that the first approach can be aligned with the forceful
motivations we listed in the previous section to introduce
indeterminate credal states, as long as we do not require that a sharp
credence represent the evidence as well as the epistemic state of
uncertainty in the agent. We have learned that this requirement can be
reduced ad absurdum even for indeterminate credal states.

One of Joyce's complaints is that a sharp credence of $0.5$ for a
\textsc{coin} contains too much information if there is little or no
evidence that the \textsc{coin} is fair. This complaint, of course, is
only effective if we make a credence say something about the evidence.
Joyce himself, however, admits that indeterminate credal states cannot
represent the evidence without violating the reflection principle due
to White's dilation problem. He is quite clear that the same
indeterminate credal state can represent different evidential
scenarios (see, for example, \scite{8}{joyce10}{302}).

In any case, Walley's and Joyce's claim that indeterminate credal
states are less informative than sharp credences has no foundation in
information theory (see \scite{8}{walley91}{34}; and
\scite{8}{joyce10}{311} for examples, but this attitude is passim). To
compare indeterminate credal states and sharp credences
informationally, we would need a non-additive set function obeying
Shannon's axioms for information. This is a non-trivial task. I have
not succeeded solving it, but I am not at all convinced that it will
result in an information measure which assigns more information to a
sharp credence such as $\{0.5\}$ than to an indeterminate credal state
such as $\{x|1/3\leq{}x\leq{}2/3\}$ for a binomial random variable.

Augustin recognizes the problem of inadequate representation long
before Joyce, with specific reference to indeterminate credal states:
\qeins{The imprecise posterior does no longer contain all the relevant
  information to produce optimal decisions. Inference and decision do
  not coincide any more} \scite{2}{augustin03}{41} (see also an
example for inadequate representation of evidence by indeterminate
credal states in \scite{8}{bradleysteele13}{16}). Indeterminate credal
states fare no better than sharp credences, except perhaps for the
problem that they unhelpfully mimic saying something about the
evidence that is much better said elsewhere.

Not only can we align sharp credences with the motivations to
introduce indeterminate credal states, we can also show that
indeterminate credal states perform worse semantically because they
mix evidential and epistemic metaphors in deleterious ways. Sharp
credences have one task: to represent epistemic uncertainty and serve
as a tool for updating, inference, and decision-making. They cannot
fulfill this task without continued reference to the evidence which
operates in the background. To use an analogy, credences are not
sufficient statistics with respect to updating, inference, and
decision-making. What is remarkable about Joyce's response to White's
dilation problem is that Joyce recognizes that indeterminate credal
states are not sufficient statistics either. But this means that they
fail at the double task which has been imposed on them: to represent
both epistemic uncertainty and the evidence.

In the following, I will provide a few examples where it becomes clear
that indeterminate credal states have difficulty representing
uncertainty because they are tangled in a double task which they
cannot fulfill.

\begin{quotex}
  \textbf{Example 1: Aggregating Expert Opinion} You have no
  information whether it will rain tomorrow ($R$) or not except the
  predictions of two weather forecasters. One of them forecasts 0.3 on
  channel GPY, the other 0.6 on channel QCT. You consider the QCT
  forecaster to be significantly more reliable, based on past
  experience.
\end{quotex}

An indeterminate credal state corresponding to this situation may be
$[0.3,0.6]$ (see \scite{8}{walley91}{214}), but it will have a
difficult time representing the difference in reliability of the
experts. A sharp credence of $P(R)=0.55$, for example, does the right
thing. Such a credence says nothing about any beliefs that the
objective chance of $R$ is $x\in{}I$ or restricted to $X\subseteq{}I$
(where $I$ is the unit interval), but it accurately reflects the
degree of uncertainty that the rational agent has over the various
possibilities. Beliefs about objective chances make little sense in
many situations where we have credences, since it is doubtful even in
the case of rain tomorrow that there is an urn of nature from which
balls are drawn. What is really at play is a complex interaction
between epistemic states (for example, experts evaluating
meteorological data) and the evidence which influences them. 

A sharp credence is often associated with distributions over chances
or other partitions of the event space, while an indeterminate credal
state puts chances in sets where they all have an equal voice. This
may also be at the bottom of Susanna Rinard's objection (see
\scite{8}{white10}{184}) that Joyce's committee members are all
equally enfranchised and so it is not clear how extremists among them
could not always be replaced by even greater extremists even after
updating on evidence which should serve to consolidate indeterminacy.
Joyce has a satisfactory response to this objection (see
\scite{8}{joyce10}{291}), but I do not see how the response addresses
the problem of aggregating expert opinion without the kind of
summation that Laplaceans find unobjectionable, even though
information is lost and can only be recouped by going back to the
evidence. More generally, the two levels for sharp credences,
representation of uncertainty and distributions over partitions,
tidily differentiate between the epistemic and the evidential
dimension; indeterminate credal states, on the other hand, just add
another level of uncertainty on top of the uncertainty that is already
expressed in the partial belief and thus do not make the appropriate
semantic distinctions.

\begin{quotex}
  \textbf{Example 2: How Precise Can a Rational Agent Reasonably Be}
  Your sharp credence for rain tomorrow, based on the expert opinion
  of channel GPY and channel QCT (you have no other information) is
  $0.55$. Is it reasonable, considering how little evidence you have,
  to reject the belief that the chance of rain tomorrow is $0.54$ or
  $0.56$; or to prefer a 54.9 cent bet on rain to a 45.1 cent bet on
  no rain?
\end{quotex}

The first question, of course, is confused, but in instructive ways (a
display of this confusion is found in \scite{8}{hajeksmithson12}{38f},
and their doctor and time of the day analogy). A sharp credence
rejects no hypothesis about objective chances (unlike an indeterminate
credal state). It often has a subjective probability distribution
operating in the background, over which it integrates to yield the
sharp credence (it would do likewise in H{\'a}jek and Smithson's
example for the prognosis of the doctor or the time of the day,
without any problems). This subjective probability distribution may
look like this:

\begin{tabular}{|lcr|} \hline
$P(\pi(R)=0.00)$ & = & $0.0001$ \\ \hline
$P(\pi(R)=0.01)$ & = & $0.0003$ \\ \hline
$P(\pi(R)=0.02)$ & = & $0.0007$ \\ \hline
$\ldots$ & & $\ldots$ \\ \hline
$P(\pi(R)=0.30)$ & = & $0.0015$ \\ \hline
$P(\pi(R)=0.31)$ & = & $0.0016$ \\ \hline
$\ldots$ & & $\ldots$ \\ \hline
$P(\pi(R)=0.54)$ & = & $0.031$ \\ \hline
$P(\pi(R)=0.55)$ & = & $0.032$ \\ \hline
$P(\pi(R)=0.56)$ & = & $0.054$ \\ \hline
$\ldots$ & & $\ldots$ \\ \hline
\end{tabular}

It is condensed by Lewis' summation formula to a sharp credence,
without being reduced to it:

\begin{equation}
  \label{eq:s2}
  P(R)=\int_{0}^{1}\zeta{}P(\pi(R)=\zeta)\,d\zeta
\end{equation}

There are more semantic questions here: what is $\pi(R)$, the
objective chance that it rains tomorrow, and how do we get to use
$P(\pi(R)=\zeta)$ in our calculation of $P(R)$ without begging the
question. Lewis' 1981 paper \qeins{A Subjectivist's Guide to
  Objective Chance} remains the gold standard in addressing these
questions, and I will no longer pursue them here. The point is that we
have properly separated the semantic dimensions and that the Laplacean
approach is not a second order probability approach. The partial
belief epistemology deals with sharp credences and how they represent
uncertainty and serve as a tool in inference, updating, and decision
making; while Lewis' Humean speculations and his interpretation of
the principal principle cover the relationship between subjective
probabilities and objective chance.

Indeterminate credal states, by contrast, mix these semantic
dimensions so that in the end we get a muddle where a superficial
reading of indeterminacy suddenly follows a converse principal
principle of sorts, namely that objective chances are constrained by
the factivity of a rational agent's credence when this credence is
knowledge (Lewis actually talks about such a converse, but in
completely different and epistemologically more intelligible terms,
see \scite{8}{lewis81}{289}; for credences being based on knowledge
see \scite{8}{levi81}{540}). Sharp credences are more, not less,
permissive with respect to objective chances operating externally
(compared to the internal epistemic state of the agent, which the
credence reflects).

The second question is also instructive: why would we prefer a 54.9
cent bet on rain to a 45.1 cent bet on no rain, given that we do not
possess the power of descrimination between these two bets? The answer
to this question ties in with the issue of incomplete preference
structure referenced above as motiviation (B) for indeterminate credal
states. 

The development of representation theorems beginning with Frank Ramsey
(followed by increasingly more compelling representation theorems in
\scite{7}{savage54}{}; and \scite{7}{jeffrey65}{}; and numerous other
variants in contemporary literature) puts the horse before the cart
and bases probability and utility functions of an agent on her
preferences, not the other way around. Once completeness as an axiom
for the preferences of an agent is jettisoned, indeterminacy follows
automatically. Indeterminacy may thus be a natural consequence of the
proper way to think about credences in terms of the preferences that
underlie them.

In response, preferences may very well logically and psychologically
precede an agent's probability and utility functions, but that does
not mean that we cannot inform the axioms we use for a rational
agent's preferences by undesirable consequences downstream.
Completeness may sound like an unreasonable imposition at the outset,
but if incompleteness has unwelcome semantic consequences for
credences, it is not illegitimate to revisit the issue. Timothy
Williamson goes through this exercise with vague concepts, showing
that all upstream logical solutions to the problem fail and that it
has to be solved downstream with an epistemic solution (see
\scite{7}{williamson96}{}). Vague concepts, like sharp credences, are
sharply bounded, but not in a way that is luminous to the agent (for
anti-luminosity see chapter 4 in \scite{7}{williamson00}{}).
Anti-luminosity answers the original question: the rational agent
prefers the 54.9 cent bet on rain to a 45.1 cent bet on no rain based
on her sharp credence without being in a position to have this
preference necessarily or have it based on physical or psychological
ability (for the analogous claim about knowledge see
\scite{8}{williamson00}{95}).

In a way, advocates of indeterminacy have solved this problem for us.
There is strong agreement among most of them that the issue of
determinacy for credences is not an issue of elicitation. The appeal
of preferences is that we can elicit them more easily than assessments
of probability and utility functions. The indeterminacy issue has been
raised to the probability level (or moved downstream) by indeterminacy
advocates themselves who feel justifiably uncomfortable with an
interpretation of their theory in behaviourist terms. So it shall be
solved there, and this paper makes an appeal to reject indeterminacy
on this level. Ironically, Isaac Levi seems to agree with me on this
point: when he talks about indeterminacy, it proceeds from the level
of probability judgment to preferences, not the other way around (see
\scite{8}{levi81}{533}).

\begin{quotex}
  \textbf{Example 3: Jaynes' Monkeys} E.T. Jaynes describes an
  experiment with monkeys filling an urn randomly with balls from
  another urn, for which sampling provides no information and so makes
  updating vacuous (see \scite{8}{jaynesbretthorst03}{160}). Here is a
  variant of this experiment for which a sharp credence provides a
  more compelling result than the associated indeterminate credal
  state: Let urn $A$ contain 4 balls, two red and two black. A monkey
  randomly fills urn $B$ from urn $A$ with two balls. We draw from urn
  $B$.
\end{quotex}

The sharp credence of drawing a red ball is $0.5$, following Lewis'
summation formula for the different combinations of balls in urn $B$.
I find this solution more intuitive in terms of further inference,
decision making, and betting behaviour than a credal state of
$\{0,1/2,1\}$, since this indeterminate credal state would licence an
exorbitant bet in favour of one colour, for example one that costs
\$9,999 and pays \$10,000 if red is drawn and nothing if black is
drawn.

\begin{quotex}
  \textbf{Example 4: Three Prisoners} Prisoner $X_{1}$ knows that two
  out of three prisoners ($X_{1},X_{2},X_{3}$) will be executed and
  one of them pardoned. He asks the warden of the prison to tell him
  the name of another prisoner who will be executed, hoping to gain
  knowledge about his own fate. When the warden tells him that $X_{3}$
  will be executed, $X_{1}$ erroneously updates his probability of
  pardon from $1/3$ to $1/2$, since either $X_{1}$ or $X_{2}$ will be
  spared.
\end{quotex}

Peter Walley maintains that for the Monty Hall problem and the Three
Prisoners problem, the probabilities of a rational agent should dilate
rather than settle on the commonly accepted solutions. Consider the
three prisoners problem. There is a compelling case for standard
conditioning and the result that the chances of pardon for prisoner
$X_{1}$ are unchanged after the update (see
\scite{8}{lukits14}{1421f}). Walley's dilated solution would give
prisoner $X_{1}$ hope on the doubtful possibility (and unfounded
assumption) that the warden might prefer to provide $X_{3}$'s name in
case prisoner $X_{1}$ was pardoned.

This example brings an interesting issue to the forefront. Sharp
credences often reflect independence of variables where such
independence is unwarranted. Booleans (more specifically, detractors
of the principle of indifference or the principle of maximum entropy,
principles which are used to generate sharp credences for rational
agents) tend to point this out gleefully. They prefer to dilate over
the possible dependence relationships (independence included). White's
dilation problem is an instance of this. The fallacy in the argument
for indeterminate credal states, illustrated by the three prisoners
problem, is that the probabilistic independence of sharp credences
does not imply independence of variables (the converse is correct),
but only that it is unknown whether there is dependence, and if yes,
whether it is correlation or inverse correlation.

\begin{quotex}
  \textbf{Example 5: Wagner's Linguist} A linguist hears the utterance
  of a native and concludes that the native cannot be part of certain
  population groups, depending on what the utterance means. The
  linguist is uncertain between some options about the meaning of the
  utterance. (For full details see \scite{8}{wagner92}{252}; and
  \scite{8}{spohn12}{197}.)
\end{quotex}

The mathematician Carl Wagner proposed a natural generalization of
Jeffrey Conditioning for his Linguist example (see
\scite{7}{wagner92}{}). Since the principle of maximum entropy is
already a generalization of Jeffrey Conditioning, the question
naturally arises whether the two generalizations agree. Wagner makes
the case that they do not agree and deduces that the principle of
maximum entropy is sometimes an inappropriate updating mechanism, in
line with many earlier criticisms of the principle of maximum entropy
(see \scite{7}{fraassen81}{}; \scite{7}{shimony85}{};
\scite{7}{skyrms87updating}{}; and, later on,
\scite{7}{grovehalpern97}{}). What is interesting about this case is
that Wagner uses indeterminate credal states for his deduction, so
that even if you agree with his natural generalization of Jeffrey
Conditioning (which I find plausible), the inconsistency with the
principle of maximum entropy can only be inferred assuming
indeterminate credal states. Wagner is unaware of this, and I am
showing in another paper (in process) how on the assumption of sharp
credences Wagner's generalization of Jeffrey conditioning perfectly
accords with the principle of maximum entropy.

This will not convince a proponent of indeterminate credences, since
they are already unlikely to believe in the general applicability of
the principle of maximum entropy (just as Wagner's argument is
unlikely to convince a proponent of the principle of maximum entropy,
since they are more likely to reject indeterminate credal states). The
battle lines are clearly drawn. Wagner's argument, instead of
undermining the principle of maximum entropy, just shows that
indeterminate credal states are as wedded to rejecting the claims of
the principle of maximum entropy as the principle of maximum entropy
is wedded to sharp credences.

Endorsement of indeterminate credal states, however, implies that
there are situations of probability update in which the posterior
probability distribution is more informative than it might be in terms
of informative theory. Indeterminate credences violate the relatively
natural intuition that we should not gain information from evidence
when a less informative updated probability will do the job of
responding to the evidence. This is not a strong argument in favour of
sharp credences. The principle of maximum of entropy has received a
thorough bashing in the last thirty years. I consider it to be much
easier to convince someone to reject indeterminate credal states on
independent (semantic) grounds than to convince them to give the
principle of maximum entropy a second chance. But the section on
semantics comes to an end here, and we want to proceed to the
intriguing issue of who does better in betting situations:
indeterminate credal states or sharp credences.

\section{Evidence Differentials and Cushioning Credences} 
\label{WalleysWorldCupWoes}

I have given away the answer already in the introduction:
indeterminate credal states do better. After I found out that agents
with indeterminate credal states do better betting on soccer games, I
let player $X$ (who uses sharp credences) and player $Y$ (who uses
indeterminate credal states) play a more basic betting game. An
$n$-sided die is rolled (by the computer). The die is fair,
unbeknownst to the players. Their bets are randomly and uniformly
drawn from the simplex for which the probabilities attributed to the
$n$ results add up to 1. Player $Y$ also surrounds her credences with
an imprecision uniformly drawn from the interval $(0,y)$. I used
Walley's pay off scheme (see \scite{8}{walley91}{632}) to settle the
bets.

The simulation results show that player $Y$ does better for $n>2$
while player $X$ does better for $n=2$. We should get similar results
if we do this analytically instead of using computer simulation. I
will pursue this further for the final version of the paper. The math
is not complicated, but unwieldy. The following expression yields the
expected gain for player $X$:

\begin{eqnarray}
  \label{eq:s3}
EX =
\frac{p_{x}}{n}\left(\sum_{j=0}^{n-1}\int_{0}^{1}\int_{0}^{x}\int_{0}^{x-y}\sum_{k=0}^{n-1}g(x,y,s,k,j)\,ds\,d\upsilon(y)\,d\xi(x)\right)+
\notag \\
\frac{p_{y}}{n}\left(\sum_{j=0}^{n-1}\int_{0}^{1}\int_{0}^{1}\int_{0}^{y-s}\sum_{k=0}^{n-1}g(x,y,s,k,j)\,d\xi(x)\,d\upsilon(y)\,ds\right)
\end{eqnarray}

where $p_{x}$ and $p_{y}$ are the respective probabilities that $X$
and $Y$ win a bet and $g$ is $X$'s gain given $X$'s credence $x$ and
$Y$'s credal state $y\pm{}s$ on roll $j$, given result $k$. $\xi(x)$
and $\upsilon(y)$ are the distributions of the credences given our
method of simplex point picking (for these distributions, one must use
the Cayley-Menger Determinant to find out the volume of generalized
pentatopes involved). A defence of sharp credences for rational agents
needs to have an explanation why for $n>2$, player $Y$ does better. We
will call it partial belief cushioning, which is based on an evidence
differential between the bettors.

In many decision-making context, we do not have the luxury of calling
off the bet. We have to decide one way or another. This is a problem
for indeterminate credal states, as Booleans have to find a way to
decide without receiving clear instructions from the credal state.
Booleans have addressed this point extensively (see for example
\scite{8}{joyce10}{311ff}; for an opponent's view of this see
\scite{8}{elga10}{6ff}). The problem for sharp credences arises when
bets are noncompulsory, for then the data above suggest that agents
holding indeterminate credal states sometimes do better. Often,
decision-making happens as betting vis-{\`a}-vis uninformed nature or
opponents which are at least as uninformed as the rational agent.
Sometimes, however, bets are offered by better informed or potentially
better informed bookies. In this case, even an agent with sharp
credences must cushion her credences and is better off by rejecting
bets that look attractive in terms of her partial beliefs. If an agent
does not cushion her partial beliefs (whether they are sharp or
indeterminate), she will incur a loss in the long run. Since
cushioning is permitted in Walley's experimental setup, Laplacean
agents should also have access to it and then no longer do worse than
Boolean agents.

Cushioning does not stand in the way of holding a sharp credence, even
if the evidence is dim. The evidence determines for a rational agent
the partial beliefs over possible states of the world operating in the
background. The better the evidence, the more pointed the
distributions of these partial beliefs will be and the more willing
the rational agent will be to enter a bet. The mathematical decision
rule will be based on the underlying distribution of the partial
beliefs, not only on the sharp credence. As we have stated before, the
sharp credence is not a sufficient statistic for decision-making,
inference, or betting behaviour; and neither is an indeterminate
credal state. If a rational agent perceives an evidence differential
and lends some belief to the proposition that the bet is offered by
someone who has more evidence about the outcome of an event than she
does, then it is likely that the rational agent will update her sharp
credence, as she would do if she were informed of another source of
expert opinion. She will certainly not be willing to enter a bet based
on her outdated sharp credence.

The rational agent with a sharp credence has resources at her disposal
to use just as much differentiation with respect to accepting and
rejecting bets as the agent with indeterminate credal states. Often
(if she is able to and especially if the bets are offered to her by a
better-informed agent), she will reject both of two complementary
bets, even when they are fair. On the one hand, any advantage that the
agent with an indeterminate credal state has over her can be
counteracted based on her distribution over partial beliefs that she
has with respect to all possibilities. On the other hand, the agent
with indeterminate credal states suffers from a semantic mixing of
metaphors between evidential and epistemic dimensions that puts her at
a real disadvantage in terms of understanding the sources and
consequences of her knowledge and her uncertainties.

\section{References} 
\label{References}

% \nocite{*} 
\bibliographystyle{ChicagoReedweb} 
\bibliography{bib-7293}

\end{document} 
